Contextual Combinatorial Cascading Bandits

### 2015-12-07 contextual combinatorial cascading bandits.tex

@@

To state our main theorem, we need some definitions. Let
$$
f_{\wedge}^* = \max_{1 \leq t \leq T} f_{\wedge, t}^{\ast}
$$
denote the best expected reward of all time under the conjunctive objective.
Let $p_{t, A}$ (resp., $p_{\wedge, t, A}$) be the probability that all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ have been observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is 
\begin{align*}
&p_{t, A} = \prod_{k=1}^{\abs{A}-1} (1 - \theta_{\ast}^{\top} x_{t, a_k^t}),\\
&p_{\wedge, t, A} = \prod_{k=1}^{\abs{A}-1} (\theta_{\ast}^{\top} x_{t, a_k^t}).
\end{align*}
We also define $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability over all time that an action could have all base arms observed under the disjunctive objective. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm, for both the disjunctive and conjunctive objectives.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and 
$\gamma_k' = 1 - \gamma_k, \gamma'_K \leq \frac{1}{4}f_{\wedge}^{\ast}$.
Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, C$^3$-UCB for both the conjunctive and disjunctive objectives, satisfy
\begin{align}
R(T) &\le \frac{\sqrt{2}}{p^*} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))}  \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln(1 + C_\gamma T/(\lambda d)) + 2\ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{p^*} \sqrt{TK} \ln T\right),
\end{align}
\begin{align}
R_{\wedge}(T) &\le \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln(1 + C_\gamma T/(\lambda d)) + 2\ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{f_{\wedge}^{\ast}}\sqrt{TK} \ln T\right).
\end{align}
%where
%\begin{align*}
%C_\gamma = \sum_{k=1}^{K} \gamma_k^2 = \sum_{k=1}^{K} (1 - \gamma_k')^2 \in [1, K].
%\end{align*}
\end{theorem}

\wei{I do not see the purpose of putting Lemma 4.4 and 4.5 in this section. Perhaps
	they should all be moved to Section 4.3}
\shuai{I think Section 4.3 will probably be put in the appendix. So I sketch a proof of main theorem here. If all the proofs will be kept, then we don't need Lemma 4.4 and 4.5 here.}

@@

\begin{lemma}
\label{lem:prefixExist}
Let $B_k = (\ba_1^t, \ldots, \ba_k^t), k \leq \abs{A}$ be a prefix of $\bA_t$. Then there exists a prefix $\bB_t$ of $\bA_t$ such that
$$
p_{\wedge, t, \bB_t} \geq \frac{1}{2} f_{\wedge, t}^* - \gamma_{\abs{\bB_t}}, \qquad R_{\wedge}(t, \bB_t) \geq \frac{1}{2} R_{\wedge}(t, \bA_t).
$$
\end{lemma}
From this lemma, we can use the prefix $\bB_t$ to substitue $\bA_t$ because $p_{\wedge, t, \bB_t}$ is large enough and $R_{\wedge}(t, \bB_t)$ will not be too small. Then follow the above deductions, we can get a good estimate for $\EE[R_{\wedge}(t, \bA_t)|\cH_{t-1}]$ in terms of $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$. Once we have estimated $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$ and $\beta_{t-1}(\delta)$, we can get an estimate for $\EE[R_{\wedge}(t, \bA_t)|\cH_{t-1}]$, thus for $R_{\wedge}(T)$. We leave all details in the next subsection.

@@

The following lemma provides the important result on the regret at each time $t$
	in terms of feature vectors of base arms in the round.
\begin{lemma}
	\label{lem:DeltaEstimate}
	For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
	$$
	R(t,A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
	$$
	if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
	$$
	R_{\wedge}(t, A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
	$$
\end{lemma}
\begin{proof}
By Lemma \ref{lem:estimateU}, $\theta_{\ast}^{\top}x_t \leq \bU_t$. Then by Lemma \ref{lem:increasing},
$$
f_t^{\ast} = f(A_t^{\ast}, \theta_{\ast}^{\top}x_t) \leq f(A_t^{\ast}, \bU_t).
$$
By Lemma \ref{lem:estimateTech} and the definition of $\bU_t$ \eqref{eq:defU},
$$
f(A, \bU_t) \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Because $f(A_t^{\ast}, \bU_t) \leq f(A, \bU_t)$, then 
$$
f_t^{\ast} \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Therefore,
$$
R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
The result for the disjunctive case can be deduced similarly.
\end{proof}

\begin{lemma}
\label{lem:prefixRelation}
Suppose 
% $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and 
$0 \leq \gamma_1' \leq \cdots \leq \gamma_K' \leq 1$. Let $A = (a_1, ..., a_{\abs{A}})$. 
% For the conjunctive case, it holds that
% $$
% p_A \geq 1 - \frac{f(A, w)}{\gamma_{\abs{A}}}.
% $$
Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. Then for the problem with conjunctive objective and $k < \abs{A}$, we have the following properties:
\begin{itemize}
\item[(1)] $f_{\wedge}(A, w) \leq \gamma_{\abs{A}}' + (1 - \gamma_{\abs{A}}') p_{\wedge, A}$;
\item[(2)] $f_{\wedge}(B_{k+1}, w) \leq f_{\wedge}(B_k, w)$;
\item[(3)] $p_{\wedge, B_{k+1}} \leq p_{\wedge, B_k}$;
\item[(4)] $f_{\wedge}(B_k, w) \leq \gamma_{k+1}' + (1 - \gamma_{k+1}') p_{B_{k+1}}$.
\end{itemize}
\end{lemma}

@@

\begin{lemma}
Suppose $0 \leq \gamma_1' \leq \cdots \leq \gamma_K' \leq 1$. Let $A = (a_1, ..., a_{\abs{A}})$. For the time $t$ and the conjunctive objective, there exists a prefix $B$ of $A$ such that 
$$
p_{\wedge, B} \geq \frac{1}{2}f_{\wedge, t}^{\ast} - \gamma_{\abs{B}}', \qquad R_{\wedge}(t, B) \geq \frac{1}{2} R_{\wedge}(t, A).
$$ 
\end{lemma}
\begin{proof}
If $f_{\wedge}(A, w) \geq \frac{1}{2} f_{\wedge, t}^{\ast}$, then by Lemma \ref{lem:prefixRelation},
$$
p_{\wedge, A} \geq \frac{1}{2}f_{\wedge, t}^{\ast} - \gamma_{\abs{A}}'.
$$

@@

\begin{proposition}
Let $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and $\gamma_k' = 1 - \gamma_k$. Suppose the equation (\ref{eq:estimateTheta}) holds for all time $t$. Then
$$
E[R(t, \bA_t)|\cH_t] \leq \frac{1}{p^*} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}
$$
If we have
% $$
% \gamma_K \geq (1+\frac{1}{n_0})f^{\ast}, \qquad 
% $$
$$
\gamma_K' \le \frac{1}{4} f_{\wedge}^{\ast},
$$
then
\begin{align*}
% &E[\Delta_{t, \bA_t}|\cH_t] \leq (n_0+1) \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},\\
&E[R_{\wedge}(t, \bA_t)|\cH_t] \leq \frac{8}{f_{\wedge}^{\ast}} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},
\end{align*}
where $f^{\ast} = \max_{t} f_t^{\ast}, f_{\wedge}^{\ast} = \max_{t} f_{\wedge, t}^{\ast}$.
\end{proposition}
\begin{proof}
For the disjunctive case,  $p_{\bA_t} \geq p^*$ by definition and the result can be derived directly.
% $p_{\bA_t} \geq \frac{1}{n_0 + 1}$ by lemma \ref{lem:prefixRelation} and the result can be derived directly.

>>>>

\newcommand{\hidetext}[1]{}

@@

To state our main theorem, we need some definitions. Let
$$
f_{\wedge}^* = \max_{1 \leq t \leq T} f_{\wedge, t}^{\ast}
$$
denote the best expected reward of all time under the conjunctive objective.
Let $p_{t, A}$ (resp., $p_{\wedge, t, A}$) be the {\em probability of full
	observation of $A$, that is the probability that all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ are observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is 
\begin{align*}
&p_{t, A} = \prod_{k=1}^{\abs{A}-1} (1 - \theta_{\ast}^{\top} x_{t, a_k^t}),\\
&p_{\wedge, t, A} = \prod_{k=1}^{\abs{A}-1} (\theta_{\ast}^{\top} x_{t, a_k^t}).
\end{align*}
We also define $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability over all time that an action could have all base arms observed under the disjunctive objective. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm, for both the disjunctive and conjunctive objectives.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and 
$\gamma_k' = 1 - \gamma_k, \gamma'_K \leq \frac{1}{4}f_{\wedge}^{\ast}$.
Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, C$^3$-UCB for both the conjunctive and disjunctive objectives, satisfy
\begin{align}
R(T) &\le \frac{\sqrt{2}}{p^*} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))}  \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln(1 + C_\gamma T/(\lambda d)) + 2\ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{p^*} \sqrt{TK} \ln T\right),
\end{align}
\begin{align}
R_{\wedge}(T) &\le \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln(1 + C_\gamma T/(\lambda d)) + 2\ln\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{f_{\wedge}^{\ast}}\sqrt{TK} \ln T\right).
\end{align}
%where
%\begin{align*}
%C_\gamma = \sum_{k=1}^{K} \gamma_k^2 = \sum_{k=1}^{K} (1 - \gamma_k')^2 \in [1, K].
%\end{align*}
\end{theorem}

\wei{I hided some paragraphs trying to explain the proof here.
	We will re-write a proof outline after revising the proofs.}
\hidetext{
\wei{I do not see the purpose of putting Lemma 4.4 and 4.5 in this section. Perhaps
	they should all be moved to Section 4.3}
\shuai{I think Section 4.3 will probably be put in the appendix. So I sketch a proof of main theorem here. If all the proofs will be kept, then we don't need Lemma 4.4 and 4.5 here.}

@@

%\begin{lemma}
%\label{lem:prefixExist}
%Let $B_k = (\ba_1^t, \ldots, \ba_k^t), k \leq \abs{A}$ be a prefix of $\bA_t$. Then there exists a prefix $\bB_t$ of $\bA_t$ such that
%$$
%p_{\wedge, t, \bB_t} \geq \frac{1}{2} f_{\wedge, t}^* - \gamma_{\abs{\bB_t}}, \qquad R_{\wedge}(t, \bB_t) \geq \frac{1}{2} R_{\wedge}(t, \bA_t).
%$$
%\end{lemma}
From this lemma, we can use the prefix $\bB_t$ to substitue $\bA_t$ because $p_{\wedge, t, \bB_t}$ is large enough and $R_{\wedge}(t, \bB_t)$ will not be too small. Then follow the above deductions, we can get a good estimate for $\EE[R_{\wedge}(t, \bA_t)|\cH_{t-1}]$ in terms of $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$. Once we have estimated $\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$ and $\beta_{t-1}(\delta)$, we can get an estimate for $\EE[R_{\wedge}(t, \bA_t)|\cH_{t-1}]$, thus for $R_{\wedge}(T)$. We leave all details in the next subsection.
}

@@

The following lemma provides the important result on the regret at each time $t$
	in terms of feature vectors of base arms at the time.
\begin{lemma}
	\label{lem:DeltaEstimate}
	For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
	$$
	R(t,A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
	$$
	if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
	$$
	R_{\wedge}(t, A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
	$$
\end{lemma}
\begin{proof}
By Lemma \ref{lem:estimateU}, $\theta_{\ast}^{\top}x_t \leq \bU_t$. Then by Lemma \ref{lem:increasing},
$$
f_t^{\ast} = f(A_t^{\ast}, \theta_{\ast}^{\top}x_t) \leq f(A_t^{\ast}, \bU_t).
$$
By Lemma \ref{lem:estimateTech} and the definition of $\bU_t$ in Eq.\eqref{eq:defU},
$$
f(A, \bU_t) \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
By our condition that $f(A_t^{\ast}, \bU_t) \leq f(A, \bU_t)$, we have 
$$
f_t^{\ast} \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
Therefore,
$$
R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t) \leq \sum_{k=1}^{\abs{A}} \gamma_k \beta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
$$
The result for the conjunctive case can be deduced in a similar way.
\end{proof}

The next lemma provides some properties about a prefix of action $A$
	in the conjunctive case, 
	which leads to the finding of a prefix $B$ of $A$ with similar regret
	and probability of full observation as those of $A$, as given
	in Lemma~\ref{lem:prefixexists}.
\begin{lemma}
\label{lem:prefixRelation}
Suppose 
% $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and 
$0 \leq \gamma_1' \leq \cdots \leq \gamma_K' \leq 1$. Let $A = (a_1, ..., a_{\abs{A}})$. 
% For the conjunctive case, it holds that
% $$
% p_A \geq 1 - \frac{f(A, w)}{\gamma_{\abs{A}}}.
% $$
Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. Then for the problem with conjunctive objective and $k < \abs{A}$, we have the following properties:
\wei{$p_{\wedge, A}$ is not defined before. Only $p_{\wedge, t, A}$ is defined.
	I guess you are using $w$ vector to be the mean vectors for $A$. Please define
	this clearly.}
\begin{itemize}
\item[(1)] $f_{\wedge}(A, w) \leq \gamma_{\abs{A}}' + (1 - \gamma_{\abs{A}}') p_{\wedge, A}$;
\item[(2)] $f_{\wedge}(B_{k+1}, w) \leq f_{\wedge}(B_k, w)$;
\item[(3)] $p_{\wedge, B_{k+1}} \leq p_{\wedge, B_k}$;
\item[(4)] $f_{\wedge}(B_k, w) \leq \gamma_{k+1}' + (1 - \gamma_{k+1}') p_{B_{k+1}}$.
\end{itemize}
\end{lemma}

@@

\begin{lemma} \label{lem:prefixexists}
Suppose $0 \leq \gamma_1' \leq \cdots \leq \gamma_K' \leq 1$. Let $A = (a_1, ..., a_{\abs{A}})$. For the time $t$ and the conjunctive objective, there exists a prefix $B$ of $A$ such that 
$$
p_{\wedge, B} \geq \frac{1}{2}f_{\wedge, t}^{\ast} - \gamma_{\abs{B}}', \qquad R_{\wedge}(t, B) \geq \frac{1}{2} R_{\wedge}(t, A).
$$ 
\end{lemma}
\begin{proof}
If $f_{\wedge}(A, w) \geq \frac{1}{2} f_{\wedge, t}^{\ast}$, then by Lemma \ref{lem:prefixRelation},
$$
p_{\wedge, A} \geq \frac{1}{2}f_{\wedge, t}^{\ast} - \gamma_{\abs{A}}'.
$$

@@

\begin{lemma}
Let $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and $\gamma_k' = 1 - \gamma_k$. Suppose the equation (\ref{eq:estimateTheta}) holds for all time $t$. Then
$$
E[R(t, \bA_t)|\cH_t] \leq \frac{1}{p^*} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}
$$
If we have
% $$
% \gamma_K \geq (1+\frac{1}{n_0})f^{\ast}, \qquad 
% $$
$$
\gamma_K' \le \frac{1}{4} f_{\wedge}^{\ast},
$$
then
\begin{align*}
% &E[\Delta_{t, \bA_t}|\cH_t] \leq (n_0+1) \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},\\
&E[R_{\wedge}(t, \bA_t)|\cH_t] \leq \frac{8}{f_{\wedge}^{\ast}} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},
\end{align*}
where $f^{\ast} = \max_{t} f_t^{\ast}, f_{\wedge}^{\ast} = \max_{t} f_{\wedge, t}^{\ast}$.
\end{lemma}
\begin{proof}
For the disjunctive case,  $p_{\bA_t} \geq p^*$ by definition and the result can be derived directly.
% $p_{\bA_t} \geq \frac{1}{n_0 + 1}$ by lemma \ref{lem:prefixRelation} and the result can be derived directly.
